Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints

### 2020-01-31 2019/NUMTA_Barkalov/SoftComputing/article.tex
Add SoftComputing

\documentclass[preprint]{elsarticle}

\usepackage{lineno,hyperref}
%\usepackage[T1]{fontenc}
\modulolinenumbers[5]

%\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints}

%% Group authors per affiliation:
\author{Roman Strongin}
\author{Konstantin Barkalov}
\author{Semen Bevzuk}
%\address[mymainaddress]{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}
\address{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia}

\begin{abstract}
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\end{abstract}

\begin{keyword}
Global optimization, Multiextremal problems, Non-convex constraints, Lipschitz constant estimates, Dimensionality reduction, Numerical methods
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}
	In this paper, we consider constrained global optimization problems and the numerical methods for solving problems of this type. Such problems are often encountered in applications (see, for example, the review \cite{Pinter2006}). As a rule, problem functions are defined by a program code, i.e. they are ``black-box'' functions, for which the computing of the values can be a time-consuming operation since in applied problems it will require  numerical modeling. Any  possibility to estimate reliably the global optimum in the multiextremal problems with ``black-box'' functions is based fundamentally on the auxiliary assumptions relating the possible values of the problem functions to their known values at the points of the trials already performed.

@@

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971, Piyavskii1972, Shubert1972, Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009, Evtushenko2013, Strongin2000, Sergeyev2013, Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems 
	\cite{Paulavicius2014, Sergeyev2015, Pinter1996, Jones1993, Gablonsky2001}
or problems with constraints of a simple structure \cite{Vaz2009, Paulavicius2016}. The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019, Pillo2012, Pillo2016}, which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints 
	\cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003} 
was used. This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions. According to the rules of the index method, every iteration called \textit{trial} at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point. When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated. This allows solving the problems with partially defined functions. The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1, Barkalov2017_2}.

	Note that the values of the Lipschitz constant for the problem functions are usually not known a priori, which  makes their estimation one of the key problems in the construction of the Lipschitzian optimization methods. The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972, Shubert1972, Wood1991, Meewella1988, Mladineo1986}) are important in the theoretical aspect, but it is difficult to use such methods in applied problems (in which the information of the Lipschitz constant values is absent).

@@

		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
\begin{itemize}[\labelitemii]
  \item global estimate of the constant $L$ in the whole search domain $D$ \cite{Strongin2000,Pinter1996,Horst1996}.
  \item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003, Sergeyev2010, Sergeyev2016}.
	\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993, Gablonsky2001,Sergeyev2006}.
\end{itemize}

@@

\section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
	We consider the global optimization problems of the form
\begin{equation}\label{problem}
	\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
	D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000, Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}%[h]
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000, Sergeyev2013})
$$
	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \; x', x''\in [0,1], \; 1\leq j \leq m+1,
$$
where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations
\begin{equation}\label{K_leq_L}
	K_j \leq 2L_j \sqrt{N+3}, \; 1\leq j \leq m+1.
\end{equation}

@@

\section{Index-Based Global Search Algorithm}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets 
$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	Determine the maximum value of the index 
$$
	M=\max\left\{\nu=\nu(x_i), \; 1 \leq i \leq k \right \}.
$$ \\
Step 3. Compute the current lower bounds
\begin{equation}\label{current_lower_bounds}
	\mu_{\nu} = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
\end{equation}
for the unknown H\"{o}lder constants $K_{\nu}$ of the functions $g_{\nu}, 1 \leq \nu \leq m+1$. If the set $I_{\nu}$ includes less than two elements or if $\mu_{\nu}$ appears to be equal to zero, then set $\mu_{\nu}=1$. \\
Step 4. For all nonempty sets $I_{\nu}, 1 \leq \nu \leq m+1$, compute the estimates
\begin{equation}\label{z_estimates}
	z_\nu^\ast = \left\{
   \begin{array}{lr}
     -\epsilon_\nu, & \nu < M,\\
     \min\{ z_i: i\in I_\nu \}, & \nu = M,
   \end{array}
	\right.
\end{equation}
where the nonnegative numbers $(\epsilon_1,\ldots,\epsilon_m )$ are the algorithm parameters. \\
Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, compute the \textit{characteristics} $R(i)$:
\begin{equation}\label{R_1}
	R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\begin{equation}\label{R_2}
	R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\end{equation}
\begin{equation}\label{R_3}
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\end{equation}
where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The values $r_{\nu} > 1, 1 \leq \nu \leq m+1$, are the parameters of the algorithm. With an appropriate selection of $r_{\nu}$ one can use the product $r_{\nu}\mu_{\nu}$ as an estimate of H\"{o}lder constants $K_{\nu}, 1 \leq \nu \leq m+1$. \\
Step 6. Find the interval $(x_{t-1}, x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
	R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation} \\
Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}, x_t$ are not the same, i.e.,
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the point
$$
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
$$
The inequality $\Delta_t \leq \epsilon$ can be taken as the termination condition, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the predefined accuracy.

@@

	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1, Barkalov2017_2}.)

\section{Index Algorithm with Dual Lipschitz Constant Estimates}
	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.  In this case, a priori estimates of the constant values are not required for the algorithm to work. These estimates are computed in the course of global search, based on the available search information.

@@

	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$: 
\begin{flushleft}
	\begin{itemize}[\labelitemii]
		\item compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$;
		\item compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$;
		\item determine the characteristic $R(i)$ as
\begin{equation}\label{R_max}
	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}},
\end{equation}
where $\rho=\Big(\frac{1-1/r_{\nu}^{glob}}{1-1/r_{\nu}^{loc}}\Big)^2$ if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$ if $\nu(x_{i-1}) \neq \nu(x_i)$.
	\end{itemize}
\end{flushleft}

@@

\section{Numerical experiments}
	To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems. In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used. This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014, Sergeyev2015,Barkalov2018}. According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as ``Simple'' and ``Hard'', respectively. The complexity of the functions is determined, in particular, by the radius of the global minimizer's attraction region and by the distance between global and local minimizers.

	In \cite{Gergel2017_2, Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
\begin{itemize}
  \item the number of constraints can be varied;
  \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
	\item the number of constraints active in the global minimizer can be predefined.
\end{itemize}

@@

\begin{table}[h!]
	\centering
		\begin{tabular}{ccccc}
			\hline
			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
			\hline
			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
			\hline
			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
			\hline
			IA & 439 & 11804(1) & 41973(3) & 404660(2) \\
			\hline
		\end{tabular}
	\caption{Average number of iterations performed by the methods on “Simple” class}
	\label{tab:average_number_of_iterations_simple}
\end{table}

>>>>

% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{lineno,hyperref}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Roman Strongin        \and
        Konstantin Barkalov   \and
				Semen Bevzuk
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}
%Insert your abstract here. Include keywords, PACS and mathematical
%subject classification numbers as needed.
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
In this paper, we consider constrained global optimization problems and the numerical methods for solving problems of this type. Such problems are often encountered in applications (see, for example, the review \cite{Pinter2006}). As a rule, problem functions are defined by a program code, i.e. they are ``black-box'' functions, for which the computing of the values can be a time-consuming operation since in applied problems it will require  numerical modeling. Any  possibility to estimate reliably the global optimum in the multiextremal problems with ``black-box'' functions is based fundamentally on the auxiliary assumptions relating the possible values of the problem functions to their known values at the points of the trials already performed.

@@

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s \cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}; since that time, this area has been continuing to develop extensively (see, for example \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems 
	\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}
or problems with constraints of a simple structure \cite{Vaz2009,Paulavicius2016}. The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method \cite{Stripinis2019,Pillo2012,Pillo2016}, which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

	Within the framework of our research, an original approach to the minimization of the multiextremal functions with non-convex constraints called the index method of the accounting for the constraints 
	\cite{Strongin2000,Pugliese,Barkalov2002,Strongin2003} 
was used. This approach is based on separate accounting for each constraint of the problem and does not involve the use of the penalty functions. According to the rules of the index method, every iteration called \textit{trial} at the corresponding point of the search domain includes a sequential checking of the feasibility of the problem constraints at this point. When the first violation of any constraint is found, the trial is terminated (the values of the remaining problem functions are not computed at this point) and the transition to the next iteration point is initiated. This allows solving the problems with partially defined functions. The experimental comparison of the index method of accounting for the constraints and the penalty function method demonstrated the advantages of the index method in terms of the number of iterations and the number of computations of the problem functions \cite{Barkalov2017_1,Barkalov2017_2}.

	Note that the values of the Lipschitz constant for the problem functions are usually not known a priori, which  makes their estimation one of the key problems in the construction of the Lipschitzian optimization methods. The methods utilizing the values of the Lipschitz constant predefined a priori (see, for example, \cite{Piyavskii1972,Shubert1972,Wood1991,Meewella1988,Mladineo1986}) are important in the theoretical aspect, but it is difficult to use such methods in applied problems (in which the information of the Lipschitz constant values is absent).

@@

		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
\begin{itemize}
\item global estimate of the constant $L$ in the whole search domain $D$ \cite{Strongin2000,Pinter1996,Horst1996}.
\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ \cite{Kvasov2003,Sergeyev2010,Sergeyev2016}.
\item choice of the estimates of the constant $L$ from some set of possible values \cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}.
\end{itemize}

@@

\section{Index Method of Accounting for the Constraints and Dimensionality Reduction}
\label{sec:1}
	We consider the global optimization problems of the form
\begin{equation}\label{problem}
	\min{\left\{\varphi(y):y \in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
	D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

@@

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{equation}\label{one_dimensional_problem}
	\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval $[0,1]$ for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

	The dimensionality reduction scheme considered here matches a multidimensional problem with Lipschitzian objective function and Lipschitzian constraints to a one-dimensional problem, in which the corresponding functions satisfy the uniform H\"{o}lder condition (see \cite{Strongin2000,Sergeyev2013})
$$
	\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x'' \right|^{1/N}, \; x', x''\in [0,1], \; 1\leq j \leq m+1,
$$
where $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related to the Lipschitz constants $L_j$ of the initial problem by the relations
\begin{equation}\label{K_leq_L}
	K_j \leq 2L_j \sqrt{N+3}, \; 1\leq j \leq m+1.
\end{equation}

@@

\section{Index-Based Global Search Algorithm}
\label{sec:2}
	The index-based global search algorithm under consideration for solving the problem (\ref{problem}) (here it is given according to \cite{Strongin2000}) involves the construction of a sequence of points $x^k \in (0,1)$, at which the search trials are conducted with the results of the type (\ref{trial_result}). The set of triplets 
$\{x^k,\; z^k=g_{\nu}\big(y(x^k)\big),\; \nu=\nu(x^k)\},\; 1 \leq k \leq n,$
makes up the search information accumulated by the method after performing $n$ steps. The operation of the index-based global search algorithm is defined by the following rules.

@@

	Determine the maximum value of the index 
$$
	M=\max\left\{\nu=\nu(x_i), \; 1 \leq i \leq k \right \}.
$$ \\
Step 3. Compute the current lower bounds
\begin{equation}\label{current_lower_bounds}
	\mu_{\nu} = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
\end{equation}
for the unknown H\"{o}lder constants $K_{\nu}$ of the functions $g_{\nu}, 1 \leq \nu \leq m+1$. If the set $I_{\nu}$ includes less than two elements or if $\mu_{\nu}$ appears to be equal to zero, then set $\mu_{\nu}=1$. \\
Step 4. For all nonempty sets $I_{\nu}, 1 \leq \nu \leq m+1$, compute the estimates
\begin{equation}\label{z_estimates}
	z_\nu^\ast = \left\{
   \begin{array}{lr}
     -\epsilon_\nu, & \nu < M,\\
     \min\{ z_i: i\in I_\nu \}, & \nu = M,
   \end{array}
	\right.
\end{equation}
where the nonnegative numbers $(\epsilon_1,\ldots,\epsilon_m )$ are the algorithm parameters. \\
Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, compute the \textit{characteristics} $R(i)$:
\begin{equation}\label{R_1}
	R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\begin{equation}\label{R_2}
	R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\end{equation}
\begin{equation}\label{R_3}
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\end{equation}
where $\Delta_i=(x_i-x_{i-1})^{1/N}$. The values $r_{\nu} > 1, 1 \leq \nu \leq m+1$, are the parameters of the algorithm. With an appropriate selection of $r_{\nu}$ one can use the product $r_{\nu}\mu_{\nu}$ as an estimate of H\"{o}lder constants $K_{\nu}, 1 \leq \nu \leq m+1$. \\
Step 6. Find the interval $(x_{t-1}, x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
	R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation} 
Step 7. Execute the next trial at the middle point of the interval $(x_{t-1}, x_t)$ if the indices of the points $x_{t-1}, x_t$ are not the same, i.e.,
$$
	x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
$$
Otherwise, execute a trial at the point
$$
	x^{k+1} = \frac{x_t+x_{t-1}}{2} + \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
$$
The inequality $\Delta_t \leq \epsilon$ can be taken as the termination condition, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the predefined accuracy.

@@

	At the same time, the index scheme of accounting for the constraints also demonstrated its advantages (in the number of computed values of the problem functions) over the traditional penalty functions method in solving a wide range of test problems (see the results of the experimental comparison in \cite{Barkalov2017_1,Barkalov2017_2}.)

\section{Index Algorithm with Dual Lipschitz Constant Estimates}
\label{sec:3}
	The index global search algorithm described in the previous section (hereafter referred to as IA) is intended for solving  the multiextremal problems, in which the objective function and the constraints satisfy the Lipschitz condition.  In this case, a priori estimates of the constant values are not required for the algorithm to work. These estimates are computed in the course of global search, based on the available search information.

@@

	Step 5. For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$: 
\begin{flushleft}
	\begin{itemize}
		\item compute the global characteristics $R_{glob}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{glob}$;
		\item compute the local characteristics $R_{loc}(i)$ according to formulae (\ref{R_1}), (\ref{R_2}), (\ref{R_3}) using the value $r_{\nu}=r_{\nu}^{loc}$;
		\item determine the characteristic $R(i)$ as
\begin{equation}\label{R_max}
	R(i) = \max{\left\{ R_{glob}(i), \rho R_{loc}(i) \right\}},
\end{equation}
where $\rho=\Big(\frac{1-1/r_{\nu}^{glob}}{1-1/r_{\nu}^{loc}}\Big)^2$ if $\nu(x_{i-1})=\nu(x_i)=\nu$, and $\rho=1$ if $\nu(x_{i-1}) \neq \nu(x_i)$.
	\end{itemize}
\end{flushleft}

@@

\section{Numerical experiments}
\label{sec:4}
To demonstrate the efficiency of the proposed algorithm, let us consider the results of solving several series of multidimensional constrained global optimization problems. In order to construct the problems, the functions generated by the GKLS generator \cite{Gaviano2003} were used. This generator of multiextremal functions is often used for investigating global optimization algorithms \cite{Paulavicius2014,Sergeyev2015,Barkalov2018}. According to \cite{Gaviano2003}, we used simpler and more complex functions, which are  further denoted as ``Simple'' and ``Hard'', respectively. The complexity of the functions is determined, in particular, by the radius of the global minimizer's attraction region and by the distance between global and local minimizers.

	In \cite{Gergel2017_2,Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
\begin{itemize}
  \item the number of constraints can be varied;
  \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
	\item the number of constraints active in the global minimizer can be predefined.
\end{itemize}

@@

\begin{table}[h!]
	\centering
	\caption{Average number of iterations performed by the methods on “Simple” class}
		\begin{tabular}{ccccc}
			\hline
			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
			\hline
			IA-DL, $r_{\nu}^{loc}=1.4$ & 316 & 7403 & 40419 & 296357 \\
			\hline
			IA-DL, $r_{\nu}^{loc}=2.0$ & 350 & 8990(1) & 41230 & 360497 \\
			\hline
			IA & 439 & 11804(1) & 41973(3) & 404660(2) \\
			\hline
		\end{tabular}
	\label{tab:average_number_of_iterations_simple}
\end{table}

\begin{table}[h!]
	\centering
	\caption{Average number of iterations performed by the methods on “Hard” class}
		\begin{tabular}{ccccc}
			\hline
			& $N=2$ & $N=3$ & $N=4$ & $N=5$ \\
			\hline
			IA-DL, $r_{\nu}^{loc}=1.4$ & 766 & 9636 & 60904 & 763711 \\
			\hline
			IA-DL, $r_{\nu}^{loc}=2.0$ & 830 & 12020 & 57473 & 743874(1) \\
			\hline
			IA & 1074 & 20327 & 87310(2) & 1084990(2) \\
			\hline
		\end{tabular}
	\label{tab:average_number_of_iterations_hard}
\end{table}
